---
title: "R Package Checker"
output: 
  html_document:
    theme: flatly
params:
  package_to_check: "dplyr"
  server_dir: "//s1428a/R_Packages/R_3_6_3_Packages"
---
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../R/get_required_packages.R") # filter cleaned
source("../R/clean_available_packages.R") # custom function to clean utils::available.packages() df
source("../R/check_r_version_required.R") # check version of R is >= to min R required for packages
source("../R/get_server_packages.R")
source("../R/compare_available_server_packages.R")
source("../R/utils.R")
```

**Note:** This analysis looks at the latest version of packages available on [CRAN](https://cran.r-project.org/) and compares this to the currently installed version of R and the copies of packages currently available on a specified server.  It does not consider older, archived versions of a package or older, archived versions of dependent packages.

```{r package-filter, include=FALSE}
search_package <- params$package_to_check
search_package_url <- paste0("https://cran.r-project.org/web/packages/", search_package, "/index.html")
server_dir <- params$server_dir

if (file.exists("../outputs/all_package_requirements.csv")) {
  all_package_df <- read_csv("../outputs/all_package_requirements.csv")
} else {
  all_package_df <- get_cleaned_available_packages()
}

package_requirements <- get_required_packages(cleaned_all_package_df = all_package_df, package_name = search_package)

check_version <- check_r_version_required(package_requirements)

# Get the zip files of R packages on server as tibble
server_packages <- get_server_packages(server_dir)

package_requirements_checked <- compare_available_server_packages(package_requirements, server_packages)

package_requirements_checked <- custom_sort_package_df(search_package, package_requirements_checked, "package")

not_on_server <- package_requirements_checked %>% filter(server_status == "package not currently on server")

need_upgrading_on_server <- package_requirements_checked %>% filter(server_status == "server version need upgrading")

packages_ok_on_server <- package_requirements_checked %>% filter(server_status == "server version currently OK")
```


## Package Requirements for [`r search_package`](`r search_package_url`)

```{r all-packages, results="asis", echo=FALSE}
if (nrow(package_requirements_checked) > 1) {
  cat(paste("The following", nrow(package_requirements_checked) - 1, "package(s) are required:<br>"))
  for (i in 2:nrow(package_requirements_checked)) {
    cat(paste0("[", package_requirements_checked[i, "package"], "]", "(", package_requirements_checked[i, "package_url"], "), "))
  }
} else {
  cat(paste("no dependent packages for", search_package))
}
```

## R version check
```{r version-check, results="asis", echo=FALSE}
r_current_version <- paste0(R.Version()$major, ".", R.Version()$minor)
if (is.logical(check_version)) {
  cat(paste("Current R version", "<b>", r_current_version, "</b>", "meets the minimum R version supported for latest version of", search_package, "and dependent packages."))
} else {
  cat(paste("Warning! Current version of R,", r_current_version, "is lower than R version required for latest version of packages:", "<br>"))
  for (i in 1:nrow(check_version)) {
    row_show <- paste(check_version[i, 1], check_version[i, 2], "requires R", check_version[i, 6], check_version[i, 7])
    cat(row_show, "<br>")
  }
}
```


## Required packages not currently on server
```{r packages-not-on-server, results="asis", echo=FALSE}
if (nrow(not_on_server) > 0) {
  cat("The following packages need adding to the server:<br>")
  for (i in 1:nrow(not_on_server)) {
    cat(paste(not_on_server[[i, "package"]], not_on_server[[i, "latest_version"]], not_on_server[[i, "package_url"]], "<br>"))
  }
} else {
  cat("All packages exist on server, but some may need upgrading (see below).")
}
```



## Required packages that need upgrading from the existing version on server
```{r packages-to-upgrade, results="asis", echo=FALSE}
if (nrow(need_upgrading_on_server) > 0) {
  cat("These packages exist on the server currently, but need upgrading:<br>")
  for (i in 1:nrow(need_upgrading_on_server)) {
    the_package <- need_upgrading_on_server[[i, "package"]]
    the_server_version <- need_upgrading_on_server[[i, "server_version"]]
    need_higher_version <- all_package_df %>%
      filter((package %in% package_requirements$package) & (dep_package == the_package) & (dep_version > the_server_version)) %>%
      pull(package)
    need_higher_version <- capture.output(cat(need_higher_version, sep = ", "))
    cat(paste("<b>", the_package, "</b>", "currently on server", "<b>", the_server_version, "</b>- requires upgrading", "to latest version", need_upgrading_on_server[[i, "latest_version"]], need_upgrading_on_server[[i, "package_url"]], "<br>(upgrade required by", need_higher_version, ")<br>"))
  }
} else {
  cat("No packages need upgrading")
}
```



## Required packages that are already on server and do not need upgrading
```{r packages-are-ok, results="asis", echo=FALSE}
if (nrow(packages_ok_on_server) >= 0) {
  cat("These packages are on the server already and meet the min required version of the package:<br>")
  for (i in 1:nrow(packages_ok_on_server)) {
    cat(paste(packages_ok_on_server[i, "package"], packages_ok_on_server[i, "server_version"], "<br>"))
  }
}
```
